---
title: "Notebooks"
author: "Data Carpentry contributors"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

------------

> ### Learning Objectives
>
> * Create an R Notebook in RStudio.
> * Enter text into an R Notebook.
> * Enter and run code chunks into an R Notebook.
> * Use Preview to quickly preview a Notebook.

------------

### About R Notebooks

This is an [R Markdown Notebook](https://rmarkdown.rstudio.com/r_notebooks.html). When you execute code within the notebook, the 
results appear beneath the code.

We will talk more about R Markdown on Day 2, and all the wonderful things you can do using it to create documents, but today we 
are going to focus on using R Markdown Notebooks to create reproducible code.

Briefly, R Markdown is a simple way to create documents with formatting by using specific symbols and characters. Programs capable 
of reading R Markdown can render a nicely formatted document. This is similar to the way that HTML files are text documents with 
specific tags, which are read by specific programs (web browsers, etc.) that apply those tags to the text, interpret the 
formatting, apply it to the document, and display it. R Markdown files (and R Notebooks) have the base file extension *.Rmd. An R 
Notebook will have the line 
output: html_notebook
in the header. 

R Notebooks have a few extra features over regular R Markdown documents:

* When you execute code in a Notebook, it is done line by line, so it can be easier to check execution for errors. In a regular R 
Markdown document, code is executed all at once.

* You can *Preview* your notebook using the *Preview* button - this displays a rendered HTML copy of the contents of the editor, 
but does not actually run any code. This makes it a bit faster to check your output.

* When you save your notebook, it creates a special *.nb.html file, which is actually a package of the HTML rendering (of the 
page, with the code sections as separate image files in their own directory) and the Markdown file. This can be shared with 
others and opened like a regular fine from within RStudio.


### Text Formatting

To create a new R Markdown Notebook in RStudio, select File -> New File -> R Notebook.

You can enter text into a Notebook as you normally would into any other text document and use Markdown conventions to format it.

Enter text between the double asterisks below to format it as bold:
** **

Enter text between the underscores below to format it as italic:
_ _

Click Preview when you are finished.

Take advantage of this to document your code and data to keep it readable, understandable, and reproducible!


### Embedding Code Chunks

You can embed R code into Notebooks by using chunks, set off above and below by three reverse ticks: ``` 

Try executing this chunk by clicking the *Run* button (small green arrow) within the chunk or by placing your cursor 
inside it and pressing *Cmd+Shift+Enter*.

```{r}
a <- 17
b <- 34
a + b
```

Here's another example - run it the same way:

```{r}
plot(cars)
```

The 'cars' dataset is an example dataset built into R, and the 'plot' command is a basic R command to create a default plot of 
simple data. We'll talk more about plotting using ggplot later on in this workshop.

Add a new chunk in the space below by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*. 
Perform the following operations:
Assign your shoe size to a variable. Then:
1. Multiply this variable by 5.
2. Add 50.
3. Multiply by 20.
4. Add 1015.
5. Subtract the year you were born.
6. Display the answer.

Insert code chunk here


### Save and Preview

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button 
or press *Cmd+Shift+K* to preview the HTML file). 

Save this notebook that you have been working in, and click *Preview* to view the output.